---
title: "NY Q2 Distributed Solar - Interactive Dashboard"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = FALSE}
library(tidyverse)
library(dplyr)
library(stringr)
library(lubridate)
library(sf)
library(leaflet)
library(leafpop)
library(htmltools)
library(scales)
library(highcharter)
library(reactable)
```

```{r include = FALSE}
coned <- read.csv('raw data/ConEd2.csv') %>% 
  select(Company, Developer, Application...Job.., Substation, PV..kWAC.,
         Application.Approved.Date...Utility., Project.Complete..Y.N., Zip.Code,
         Metering..NA...NM...RNM...CDG.) #%>% 
  #mutate(date = mdy(Application.Approved.Date...Utility.))
nyseg <- read.csv('raw data/NYSEG2.csv') %>% 
  select(Company, Developer, Application...Job.., Substation, PV..kWAC.,
         Application.Approved.Date...Utility., Project.Complete..Y.N., Zip.Code,
         Metering..NA...NM...RNM...CDG.)
or <- read.csv('raw data/O^0R2.csv') %>% 
  select(Company, Developer, Application...Job.., Substation, PV..kWAC.,
         Application.Approved.Date...Utility., Project.Complete..Y.N., Zip.Code,
         Metering..NA...NM...RNM...CDG.)
nimo <- read.csv('raw data/Niagara Mohawk (NatGrid)2.csv') %>% 
  select(Company, Developer, Application...Job.., Substation, PV..kWAC.,
         Application.Approved.Date...Utility., Project.Complete..Y.N., Zip.Code,
         Metering..NA...NM...RNM...CDG.)
ch <- read.csv('raw data/CH2.csv') %>% 
  select(Company, Developer, Application...Job.., Substation, PV..kWAC.,
         Application.Approved.Date...Utility., Project.Complete..Y.N., Zip.Code,
         Metering..NA...NM...RNM...CDG.)
pseg <- read.csv('raw data/PSEG2.csv') %>% 
  select(Company, Developer, Application...Job.., Substation, PV..kWAC.,
         Application.Approved.Date...Utility., Project.Complete..Y.N., Zip.Code,
         Metering..NA...NM...RNM...CDG.)

ny2 <- rbind(coned, nimo, nyseg, or, ch, pseg)

rm(coned, nimo, nyseg, or, ch, pseg)

ny_clean2 <- ny2 %>% 
  rename('utility' = 'Company', 'developer' = 'Developer', 'id' = 'Application...Job..',
         'substation' = 'Substation', 'capacity' = 'PV..kWAC.', 'date' = 'Application.Approved.Date...Utility.',
         'complete' = 'Project.Complete..Y.N.', 'zip' = 'Zip.Code') %>% 
  #filter(!is.na(Metering..NA...NM...RNM...CDG.)) %>% 
  mutate(date = dmy(date),
         capacity = as.numeric(capacity),
         segment_x = case_when(
           Metering..NA...NM...RNM...CDG. == 'CDG' ~ 'Community Solar',
           capacity <= 20 ~ 'Residential',
           capacity > 20 ~ 'Non-Residential'
         ),
         n = 1,
         quarter = zoo::as.yearqtr(date, format = "%Y-%m-%d")) %>% 
  filter(!is.na(capacity))
```

```{r include = FALSE}
ny_sum <- ny_clean2 %>% 
  filter(date >= '2020-07-01',
         quarter != '2021 Q3') %>% 
  group_by(segment_x, quarter) %>% 
  summarise(cap = sum(capacity)/1000)

ny_2 <- ny_clean2 %>% 
  filter(quarter == '2021 Q2') %>% 
  mutate(res_cap = if_else(segment_x == 'Residential', capacity, 0),
         com_cap = if_else(segment_x == 'Non-Residential', capacity, 0),
         cs_cap = if_else(segment_x == 'Community Solar', capacity, 0),
        # month = months(as.Date(date)))
        month = month(date))
```

## New York added **`r round(sum(ny_2$capacity)/1000,2)` MW** of new distributed solar capacity to the interconnection queue in Q2 2021

```{r echo = FALSE}
ny_sum %>%
  hchart('column', hcaes(x = as.factor(quarter), y = round(cap,2), group = segment_x)) %>% 
  hc_colors(c('#F37325', '#2490BD', "#47B970", '#1A2B40', '#b7b7b7', '#800000','#F8AA1A')) %>% 
  hc_plotOptions(series = list(stacking = 'normal')) %>% 
  hc_xAxis(title = '', labels = list(step = 1) #, minorTickInterval
  ) %>% 
  hc_yAxis(title = list(text = 'Quarterly capacity (MW)'), style = list(fontSize = "5.0vh"))
       
```

```{r include = FALSE}
ny_2_seg2 <- ny_2 %>% 
  mutate(meter = Metering..NA...NM...RNM...CDG.) %>% 
  mutate(meter = as.factor(case_when(
    is.na(meter) | meter == 'n/a' ~ 'Non-Metered',
    meter == 'NM' ~ 'Net Metered',
    meter == 'RNM' ~ 'Remote Net Metered',
    meter == 'CDG' ~ 'Community Solar'
  ))) %>% 
  group_by(meter) %>% 
  summarise(cap = sum(capacity),
            n = n(),
            mean = mean(capacity)) %>% 
  ungroup() %>% 
  mutate(share = cap/sum(cap))
```

## Of this **`r round(sum(ny_2$capacity/1000),2)` MW** of new capacity...

```{r results = 'asis', echo = FALSE}
for (k in levels(ny_2_seg2$meter)) {
  cat("- ", "**", round(ny_2_seg2$cap[ny_2_seg2$meter == k]/1000,2), "**", " MW was ", k, sep = "")
  cat("\n")
}
```

```{r echo = FALSE}
ny_2_seg2 %>% 
  hchart('pie', hcaes(x = meter, y = round(cap/1000,2)), name = "Capacity (MW)") %>% 
  hc_colors(c( '#2490BD', '#F37325', "#47B970", '#1A2B40', '#b7b7b7', '#F8AA1A', '#800000')) 

```


## Which utilities added the most capacity in Q2 2021?

```{r include = FALSE}
ny_2_ut <- ny_2 %>%
  group_by(utility) %>% 
  summarise(cap = sum(capacity))

ny_2_ut2 <- ny_2 %>%
  group_by(utility, month) %>% 
  summarise(cap = sum(capacity)) %>% 
  mutate(month = factor(month, levels = c(4, 5, 6), labels = c('April', 'May', 'June')))
```

```{r echo = FALSE}
ny_2_ut %>% 
  hchart('pie', hcaes(x = utility, y = round(cap/1000,2)), name = "Capacity (MW)") %>% 
  hc_colors(c('#F37325', '#2490BD', "#47B970", '#b7b7b7', '#1A2B40', '#800000','#F8AA1A')) 

```

```{r echo = FALSE}
ny_2_ut2 %>% 
#  mutate(month = factor(month, levels = c('April', 'May', 'June'), labels = c('April', 'May', #'June'))) %>% 
  hchart('column', hcaes(x = month, y = round(cap/1000, 2), group = utility)) %>% 
  hc_colors(c('#F37325', '#2490BD', "#47B970", '#b7b7b7', '#1A2B40', '#800000','#F8AA1A')) %>% 
  hc_plotOptions(series = list(stacking = 'normal')) %>% 
  hc_xAxis(title = '', labels = list(step = 1) #, minorTickInterval
  ) %>% 
  hc_yAxis(title = list(text = 'Monthly capacity (MW)'), style = list(fontSize = "5.0vh"))
       

```

## Which developers added the most capacity in Q2 2021?

```{r include = FALSE}
ny_2 <- ny_2 %>% 
  mutate(dev_agg = case_when(
    grepl('tesla|solarcity|solar city', developer, ignore.case = T) ~ 'Tesla/SolarCity',
    grepl('momentum', developer, ignore.case = T) ~ 'Momentum Solar',
    grepl('vivint|sunrun', developer, ignore.case = T) ~ 'Vivint/SunRun',
    grepl('sunpower|sun power', developer, ignore.case = T) ~ 'SunPower',
    grepl('venture', developer, ignore.case = T) ~ 'Venture Solar',
    grepl('trinity', developer, ignore.case = T) ~ 'Trinity Solar',
    grepl('empire', developer, ignore.case = T) ~ 'Empire Solar Solutions',
    grepl('plug', developer, ignore.case = T) ~ 'PlugPV',
    grepl('kasselman', developer, ignore.case = T) ~ 'Kasselman Solar',
    grepl('new york state solar|ny state solar', developer, ignore.case = T) ~ 'New York State Solar Farm',
    grepl('suncommon', developer, ignore.case = T) ~ 'SunCommon',
    grepl('infinity', developer, ignore.case = T) ~ 'Infinity Solar',
    grepl('borrego', developer, ignore.case = T) ~ 'Borrego Solar',
    grepl('resunance', developer, ignore.case = T) ~ 'RESunance',
    grepl('atlas', developer, ignore.case = T) ~ 'Atlas Renewables (CAERO Group)',
    grepl('amp solar d|amp solar u', developer, ignore.case = T) ~ 'Amp Solar',
    grepl('nexamp', developer, ignore.case = T) ~ 'Nexamp',
    grepl('renesola', developer, ignore.case = T) ~ 'ReneSolar Power Holdings',
    grepl('g&s', developer, ignore.case = T) ~ 'G&S Solar',
    grepl('delaware', developer, ignore.case = T) ~ 'Delaware River Solar',
    grepl('arcadia', developer, ignore.case = T) ~ 'Arcadia',
    grepl('oya', developer, ignore.case = T) ~ 'OYA Solar',
    grepl('citizens', developer, ignore.case = T) ~ 'Citizens Energy',
    grepl('long island', developer, ignore.case = T) ~ 'Long Island Power Solutions',
    grepl('kamtech|kam tech', developer, ignore.case = T) ~ 'Kamtech Solar',
    grepl('harvest', developer, ignore.case = T) ~ 'Harvest Power',
    grepl('empower', developer, ignore.case = T) ~ 'EmPower Solar',
    grepl('sunation', developer, ignore.case = T) ~ 'SUNation Solar',
    grepl('montante', developer, ignore.case = T) ~ 'Montante Solar',
    grepl('liberty', developer, ignore.case = T) ~ 'Solar Liberty Energy',
    grepl('lodestar', developer, ignore.case = T) ~ 'Lodestar Energy',
    grepl('omni', developer, ignore.case = T) ~ 'Omni-Navitas',
    grepl('bw ', developer, ignore.case = T) ~ 'BW Solar',
    grepl('aura', developer, ignore.case = T) ~ 'Aura Power',
    grepl('ric d|ric e', developer, ignore.case = T) ~ 'RIC Development',
    grepl('caero', developer, ignore.case = T) ~ 'CAERO Group',
    grepl('solitude', developer, ignore.case = T) ~ 'Solitude Solar',
    grepl('cypress c', developer, ignore.case = T) ~ 'Cypress Creek Renewables',
    grepl('saturn', developer, ignore.case = T) ~ 'Saturn Power Corporation',
    grepl('green l|greenl', developer, ignore.case = T) ~ 'Green Logic',
    grepl('hudson', developer, ignore.case = T) ~ 'Hudson River',
    grepl('microgrid', developer, ignore.case = T) ~ 'MicroGrid Networks',
    grepl('nextera|next era|dg new', developer, ignore.case = T) ~ 'NextEra Energy (DG New York CS)',
    grepl('eden', developer, ignore.case = T) ~ 'Eden Renewables',
    grepl('nexamp', developer, ignore.case = T) ~ 'Nexamp',
    grepl('source r', developer, ignore.case = T) ~ 'Source Renewables',
    grepl('CVE', developer) ~ 'CVE North America',
    grepl('convergent', developer) ~ 'Convergent Energy & Power',
    grepl('Nexus', developer) ~ 'Nexus Renewables',
    grepl('Dyna', developer) ~ 'DynaSolar DevCo',
    grepl('siemens', developer, ignore.case = T) ~ 'Siemens',
    grepl('Oriden', developer) ~ 'Oriden',
    grepl('sunrise', developer, ignore.case = T) ~ 'Sunrise Power Solutions',
    TRUE ~ developer    
  ))

ny_2_dev <- ny_2 %>% 
  group_by(dev_agg, segment_x) %>% 
  summarise(cap = sum(capacity),
            n = n(),
            mean = mean(capacity)) %>% 
  ungroup() %>% 
  mutate(share = cap/sum(cap)) %>% 
  arrange(desc(cap))
```

### Top 10 developers in the residential segment:

```{r echo = FALSE}
ny_2_dev %>% filter(segment_x == 'Residential') %>%
  head(10) %>%
  select(dev_agg, cap, n, share) %>% 
  mutate(share = paste(round(share*100,2), "%", sep = ""),
         cap = format(round(as.numeric(cap), 2), big.mark = ",")) %>% 
  knitr::kable(col.names = c("Developer", "Capacity (kW)", "Number of Projects", "Market Share"),
               digits = 2, align = 'l') %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F) %>% 
  kableExtra::add_header_above(c("Q2 2021" = 4))
```


### Top 10 developers in the commercial segment:

```{r echo = FALSE}
ny_2_dev %>% filter(segment_x == 'Non-Residential') %>%
  head(10) %>%
  select(dev_agg, cap, n, share) %>% 
  mutate(share = paste(round(share*100,2), "%", sep = ""),
         cap = format(round(as.numeric(cap), 2), big.mark = ",")) %>% 
  knitr::kable(col.names = c("Developer", "Capacity (kW)", "Number of Projects", "Market Share"),
               digits = 2, align = 'l') %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F) %>% 
  kableExtra::add_header_above(c("Q2 2021" = 4))
```

### Top 10 developers active in the community solar segment:

```{r echo = FALSE}
ny_2_dev %>% filter(segment_x == 'Community Solar') %>%
  head(10) %>%
  select(dev_agg, cap, n, share) %>% 
  mutate(share = paste(round(share*100,2), "%", sep = ""),
         cap = format(round(as.numeric(cap), 2), big.mark = ",")) %>% 
  knitr::kable(col.names = c("Developer", "Capacity (kW)", "Number of Projects", "Market Share"),
               digits = 2, align = 'l') %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F) %>% 
  kableExtra::add_header_above(c("Q2 2021" = 4))
```

## Where was new capacity added in Q2 2021?

```{r include = FALSE}
ny_map <- ny_2 %>% 
  mutate(geoid2 = as.numeric(substr(zip,1,5))) %>% 
  filter(!geoid2 == 16807|!geoid2 == 17047) %>% 
  group_by(geoid2) %>% 
  summarise(cap = sum(capacity)/1000)

nymap <- st_read('raw data/acs2019_5yr_B01003_86000US14475.shp') %>% 
  mutate(geoid2 = as.numeric(substr(geoid,8,12))) %>% 
  filter(!name == 'New York')

ny__map <- left_join(nymap, ny_map, by = 'geoid2')
```

```{r include = FALSE}
zip <- read.csv('raw data/uszips.csv')

zip_join <- zip %>% 
  select(zip, city) %>% 
  rename('geoid2' = 'zip')

ny_int <- left_join(ny__map, zip_join, by = 'geoid2') %>% 
  mutate(cap = cap*1000)

ny_int <- filter(ny_int, name != 'New Jersey')
ny_int <- mutate(ny_int, cap = round(cap,0), cap2 = if_else(is.na(cap),0,cap))

blue_mon <- c('#e0f1f9', '#2490BD') 

pal <- colorNumeric(palette = blue_mon, domain = ny_int$cap, na.color = '#FFFFFF')
labels <- sprintf("<strong>%s</strong><br/> Zip code: %s<br/> Capacity Q2 2021: %s kW",
                  ny_int$city, paste(0, ny_int$geoid2, sep = ""), comma(ny_int$cap2)) %>% 
  lapply(HTML)

map_int <- leaflet(ny_int) %>% addTiles()
```

```{r echo = FALSE}
map_int %>% addPolygons(fillColor = ~pal(cap),
                        weight = 0.5,
                        opacity = 1,
                        color = "gray",
                        dashArray = "3",
                        fillOpacity = 0.7,
                        highlightOptions = highlightOptions(color = '#2a2a2a',
                                                            weight = 2,
                                                            dashArray = "",
                                                            fillOpacity = 0.7,
                                                            bringToFront = T),
                        label = labels,
                        labelOptions = labelOptions(
                          style = list("font-weight" = "normal", padding = "3px 8px"),
                          textsize = "15px",
                          direction = "auto",
                          opacity = 0.75)) %>% 
  addProviderTiles('Esri.WorldGrayCanvas')
```

## Filter and download data

```{r include = FALSE}
ny_2_table <- ny_2 %>% 
  mutate(dev_agg = case_when(
    grepl('tesla|solarcity|solar city', developer, ignore.case = T) ~ 'Tesla/SolarCity',
    grepl('momentum', developer, ignore.case = T) ~ 'Momentum Solar',
    grepl('vivint|sunrun', developer, ignore.case = T) ~ 'Vivint/SunRun',
    grepl('sunpower|sun power', developer, ignore.case = T) ~ 'SunPower',
    grepl('venture', developer, ignore.case = T) ~ 'Venture Solar',
    grepl('trinity', developer, ignore.case = T) ~ 'Trinity Solar',
    grepl('empire', developer, ignore.case = T) ~ 'Empire Solar Solutions',
    grepl('plug', developer, ignore.case = T) ~ 'PlugPV',
    grepl('kasselman', developer, ignore.case = T) ~ 'Kasselman Solar',
    grepl('new york state solar|ny state solar', developer, ignore.case = T) ~ 'New York State Solar Farm',
    grepl('suncommon', developer, ignore.case = T) ~ 'SunCommon',
    grepl('infinity', developer, ignore.case = T) ~ 'Infinity Solar',
    grepl('borrego', developer, ignore.case = T) ~ 'Borrego Solar',
    grepl('resunance', developer, ignore.case = T) ~ 'RESunance',
    grepl('atlas', developer, ignore.case = T) ~ 'Atlas Renewables (CAERO Group)',
    grepl('amp solar d|amp solar u', developer, ignore.case = T) ~ 'Amp Solar',
    grepl('nexamp', developer, ignore.case = T) ~ 'Nexamp',
    grepl('renesola', developer, ignore.case = T) ~ 'ReneSolar Power Holdings',
    grepl('g&s', developer, ignore.case = T) ~ 'G&S Solar',
    grepl('delaware', developer, ignore.case = T) ~ 'Delaware River Solar',
    grepl('arcadia', developer, ignore.case = T) ~ 'Arcadia',
    grepl('oya', developer, ignore.case = T) ~ 'OYA Solar',
    grepl('citizens', developer, ignore.case = T) ~ 'Citizens Energy',
    grepl('long island', developer, ignore.case = T) ~ 'Long Island Power Solutions',
    grepl('kamtech|kam tech', developer, ignore.case = T) ~ 'Kamtech Solar',
    grepl('harvest', developer, ignore.case = T) ~ 'Harvest Power',
    grepl('empower', developer, ignore.case = T) ~ 'EmPower Solar',
    grepl('sunation', developer, ignore.case = T) ~ 'SUNation Solar',
    grepl('montante', developer, ignore.case = T) ~ 'Montante Solar',
    grepl('liberty', developer, ignore.case = T) ~ 'Solar Liberty Energy',
    grepl('lodestar', developer, ignore.case = T) ~ 'Lodestar Energy',
    grepl('omni', developer, ignore.case = T) ~ 'Omni-Navitas',
    grepl('bw ', developer, ignore.case = T) ~ 'BW Solar',
    grepl('aura', developer, ignore.case = T) ~ 'Aura Power',
    grepl('ric d|ric e', developer, ignore.case = T) ~ 'RIC Development',
    grepl('caero', developer, ignore.case = T) ~ 'CAERO Group',
    grepl('solitude', developer, ignore.case = T) ~ 'Solitude Solar',
    grepl('cypress c', developer, ignore.case = T) ~ 'Cypress Creek Renewables',
    grepl('saturn', developer, ignore.case = T) ~ 'Saturn Power Corporation',
    grepl('green l|greenl', developer, ignore.case = T) ~ 'Green Logic',
    grepl('hudson', developer, ignore.case = T) ~ 'Hudson River',
    grepl('microgrid', developer, ignore.case = T) ~ 'MicroGrid Networks',
    grepl('nextera|next era|dg new', developer, ignore.case = T) ~ 'NextEra Energy (DG New York CS)',
    grepl('eden', developer, ignore.case = T) ~ 'Eden Renewables',
    grepl('nexamp', developer, ignore.case = T) ~ 'Nexamp',
    grepl('source r', developer, ignore.case = T) ~ 'Source Renewables',
    grepl('CVE', developer) ~ 'CVE North America',
    grepl('convergent', developer) ~ 'Convergent Energy & Power',
    grepl('Nexus', developer) ~ 'Nexus Renewables',
    grepl('Dyna', developer) ~ 'DynaSolar DevCo',
    grepl('siemens', developer, ignore.case = T) ~ 'Siemens',
    grepl('Oriden', developer) ~ 'Oriden',
    grepl('sunrise', developer, ignore.case = T) ~ 'Sunrise Power Solutions',
    TRUE ~ developer    
  )) %>% 
  select(dev_agg, capacity, date, utility, segment_x, zip) %>% 
 # mutate('Number of Projects' = 1) %>% 
  rename('Developer' = dev_agg, 'Capacity' = capacity, 'Date' = date, 'Utility' = utility, 'Segment' = segment_x, 'Zip Code' = zip)

```


```{r echo = FALSE}

reactable(ny_2_table, filterable = TRUE,
          columns = list(Capacity = colDef(filterable = FALSE)),
          searchable = TRUE,
          showPageSizeOptions = TRUE,
          pageSizeOptions = c(5,10,20,50),
          defaultPageSize = 10)

```
